---
title: "About this workshop"
---


```{r functions, echo=FALSE}
# Generate package chunk
   file_names <- stringr::str_subset(dir(), "qmd")
  packages_all <- NULL
  for(i in file_names){
  	#print(i)
  	
  lines_text <- readLines(i) 
  packages <- gsub("library\\(|\\)", "",  
                  unlist(stringr::str_extract_all(lines_text,  
                                         "library\\([a-zA-z0-9]*\\)|p_load\\([a-zA-z0-9]*\\)"))) 
  packages <- packages[packages!="pacman"] 
  packages <- packages[packages!="p_load("] 
  packages_all <- c(packages_all, packages)
  
  }
  packages <- paste("# install.packages('pacman')", "library(pacman)", "p_load('", paste(unique(packages_all), collapse="', '"), "')",sep="") 
  packages <- stringr::str_wrap(packages, width = 80) 
  packages <- gsub("install.packages\\('pacman'\\)", "install.packages\\('pacman'\\)\n", packages) 
  packages <- gsub("library\\(pacman\\)", "library\\(pacman\\)\n", packages)
```







## About us (Paul & Jonas)
- Paul
	- Currently [MZES](https://www.mzes.uni-mannheim.de/d7/en) Research fellow (University of Mannheim)
	- Previously... University of Bern (PhD, 2015), European University Institute (Postdoc, 15'-17')
	- My research interests: political sociology & methodology [see [Google Scholar](https://scholar.google.com/citations?user=hw-b-MkAAAAJ&hl=en&oi=ao))
	- Started R around 2009, first shiny apps ~2015 for teaching/research purposes (e.g., [here](https://paulcbauer.shinyapps.io/transformations_of_variables/) and [here](https://paulcbauer.shinyapps.io/visualizing-causal-scenarios/))
	- Contact: `mail@paulcbauer.de`; [www.paulcbauer.de](https://sites.google.com/paulcbauer.de/web/); [Twitter](https://twitter.com/p_c_bauer); [Github](https://github.com/paulcbauer)
- Jonas
	- Currently a research fellow at GESIS' survey data curation (SDC) department
	- Constantly learning about R and Shiny since 2020
	- Interested in social geography, spatial analysis and computational text analysis
	- Contact: `jonas.lieth@gesis.org`; GitHub: [JsLth](https://github.com/JsLth)





## Your turn

- Let's check our the survey results...

1. Name?
2. Affiliation? Country?
3. Research interests/questions?

<!--
Below take too long for short workshop
4. Previous experience with R?
5. Why are you interested in this
workshop?
-->




## Contact & Outline & Dates
- **Important**: 1st time we teach workshop/material
- **Course outline/content/dates:** (see toc on the left) 
  - Day 1: Introduction to Shiny
    - 13:00 - 14:00: Welcome and Introduction (1)
    - 14:00 - 15:00: Introduction (2)
    - 15:00 - 15:30: Coffee break
    - 15:30 - 16:30: User Interface (UI): Designing the front end (1)
    - 16:30 - 17:30: User Interface (UI): Designing the front end (2) [**Introduction tab**]
  - Day 2: Reactive programming
    - 13:00 - 14:00: Server: Reactive programming (1)
    - 14:00 - 15:00: Server: Reactive programming (2)  [**Tabulate data tab**]
    - 15:00 - 15:30: Coffee break
    - 15:30 - 16:30: Modelling and visualizing data (1)
    - 16:30 - 17:30: Modelling and visualizing data (2) [**Modelling data tab**]
  - Day 3: Develop your app
    - 13:00 - 14:00: Mapping data & advanced visualization (1)
    - 14:00 - 15:00: Mapping data & advanced visualization (2)  [**Mapping data tab**]
    - 15:00 - 15:30: Coffee break
    - 15:30 - 16:30: Theming & styling 
    - 16:30 - 17:30: Strengthen & Deploy






## Script & material {#material}
- **Literature**: See syllabus.
- **Website/script**: https://bookdown.org/paul/shiny_workshop/
	- Find it: Google "shiny paul jonas"
	- Document = slides  + script (Zoom in/out with `STRG + mousewheel`)
	- **Code**: can all be found in the script
	- **Data**: can usually be downloaded over links in the script. If not we'll share the files.
	- Full screen: F11
	- **Navigation**: TOCs on left and right
	- Search document (upper left)
	- Document generated with [quarto](https://quarto.org/)
- **Motivation**: Have a go-to script for participants (and ourselves!)
- **Content**: Mixture of theory, lab sessions, exercises and pure code examples for discussion




## Strategy & Goals
- **Strategy**: From the simple to the complex, slowly building up a complex Shiny app that includes various aspects (tabulate data, modelling and descriptive graphs, mapping)

- **Goals**: By the end of the course participants will:
	- understand the basics of reactive programming for interactive data analysis and visualization
	- know what the structure of a Shiny application looks like
	- be comfortable to use R Shiny to build their own interactive applications
	- have learned about different ways to launch their Shiny application






  






## Online vs. offline

* **Negative**
    + Screen fatigue
    + Can't run around to check your code
    + Less engaging, less social
    + Voice
    + Screen sharing &less screen space than classroom 
* **Positive**
    + We see the Shiny app how its mostly consumed
    ;-) (on a screen)
        + Remember: "How is your Shiny app consumed (smartphone)?"
    + Participation from everywhere
* **Rule(s)**: Please keep your camera online!
  + Distracting animals/children/partners are a welcome distraction!
  + Yawning, leaving, looking bored etc. allowed!
  + Use a virtual background if you like!
  







  
    
## Recommended readings

- Important: The course does not require any prior reading. 
- However, our schedule is primarily based on two textbooks which we generally recommend for further reading:
	- @wickham2021mastering: Mastering Shiny: Build Interactive Apps, Reports, and Dashboards Powered by R. Accessible online at: https://mastering-shiny.org/. 
	- Fay, Colin, Rochette, Sébastien, Guyader, Vincent, and Girard, Cervan (2022): Engineering Production-Grade Shiny Apps. Accessible online at: https://engineering-shiny.org/.













## Software we will use
- Open-source software! (Q: Why?)
- R [@rcoreteam2023]^[Creators: [Core contributors](https://www.r-project.org/contributors.html)  and thousands of package authors.]
  - only viable competitor is Python
  - Install the necessary packages using the code below.

```{r workshop-1, echo=FALSE, comment=NA}
options(width = 300)
cat(packages)
```

- Shiny [@chang2020; @wickham2021mastering]
- Ggplot2^[Creators: https://github.com/tidyverse/ggplot2] [@wickham2016]
* Plotly^[Creators: https://github.com/plotly/plotly.js; https://github.com/ropensci/plotly] [@sievert2020]
* Note: Ideally cite the software you use in your research especially when it is open-source (e.g., run `citation("ggplot2")`)






## Helpful resources

1. [Shiny user showcase](https://shiny.posit.co/r/gallery/#user-showcase) 
2. [Shiny feature demos](https://shiny.posit.co/r/gallery/#feature-demos)
3. [Shiny widget gallery](https://shiny.posit.co/r/gallery/widgets/widget-gallery/)
4. [R Shiny for Enterprise Application Development: 6 Lessons for Building Shiny Apps](https://appsilon.com/why-you-should-use-r-shiny-for-enterprise-application-development/#anchor-7)
5. [Shiny apps by Dimiter Toshkov](http://www.dimiter.eu/Visualizations.html)
