---
title: "About this workshop"
---


```{r functions, echo=FALSE, cache=FALSE}
# Manual packages
 packages_manual <- c("profvis", "rsconnect", "whereami", "DT")


# Generate package chunk
   file_names <- stringr::str_subset(dir(), "qmd")
  packages_all <- NULL
  for(i in file_names){
  	#print(i)
  	
  lines_text <- readLines(i) 
  packages <- gsub("library\\(|\\)", "",  
                  unlist(stringr::str_extract_all(lines_text,  
                                         "library\\([a-zA-z0-9]*\\)|p_load\\([a-zA-z0-9]*\\)"))) 
  packages <- packages[packages!="pacman"] 
  packages <- packages[packages!="p_load("] 
  packages_all <- c(packages_all, packages)
  
  }
  # Add manual packages
  packages_all <- c(packages_all, packages_manual)
  
  # Clean up
  packages <- paste("# install.packages('pacman')", "library(pacman)", "p_load('", paste(unique(packages_all), collapse="', '"), "')",sep="") 
  packages <- stringr::str_wrap(packages, width = 80) 
  packages <- gsub("install.packages\\('pacman'\\)", "install.packages\\('pacman'\\)\n", packages) 
  packages <- gsub("library\\(pacman\\)", "library\\(pacman\\)\n", packages)

```



This document serves as slides and script for the workshop ***Interactive Data Analysis with Shiny*** taught by [Paul C. Bauer](http://paulcbauer.eu/) <a href='https://twitter.com/p_c_bauer' class='fa fa-twitter' target='_blank'></a> and [Jonas Lieth]() (Gesis, Mannheim, Online, 5-7th of July 2023). Original material is licensed under a [Attribution-NonCommercial-ShareAlike 4.0 International (CC BY-NC-SA 4.0)](https://creativecommons.org/licenses/by-nc-sa/4.0/) license. Where we draw on other authors material other licenses may apply (see references in the syllabus as well as the citations and links in the script). For potential future versions of this material see the  [github repository](https://github.com/paulcbauer/shiny_workshop/). If you have feedback or discover errors/dead links please let us know via email (mail@paulcbauer.de/Jonas.Lieth@gesis.org) or by submitting an issue on github. And make sure to [install all the necessary packages]() before you use the script.

<br><br>

## About us (Paul & Jonas)
- Paul	
  + Instructor at University of Freiburg (Department of politics), Postdoc at LMU Munich (statistics department, Frauke Kreuter), External fellow at [MZES](https://www.mzes.uni-mannheim.de/d7/en)
  + Previously
    + PhD at the [University of Bern](http://www.unibe.ch/index_eng.html)
    + research/postdoctoral fellow at the [MZES](https://www.mzes.uni-mannheim.de/d7/en), the [EUI](https://www.eui.eu/) (Florence, Italy)
  + My [research](https://sites.google.com/view/paulcbauer/publications) ([Google Scholar](https://scholar.google.ch/citations?user=hw-b-MkAAAAJ&hl=en), [Github](https://github.com/paulcbauer))
    + **Substantive**: Political sociology & comparative politics (Trust, polarization, social media, fake news)
    + **Methods/data**: Causal inference, experiments, text data, data visualization, machine learning, big data (Google trends, Twitter)
- Jonas
	- Currently a research fellow at GESIS' survey data curation (SDC) department
	- PhD student interested in social geography, spatial data science, and computational text analysis
	- Developing R packages (e.g. [`r311`](https://github.com/JsLth/r311)) and Shiny apps  (e.g. [`gretan`](https://github.com/JsLth/gretan)) for research
	- Contact: `jonas.lieth@gesis.org`; [Website](https://www.gesis.org/institut/ueber-uns/mitarbeitendenverzeichnis/person/Jonas.Lieth); [GitHub](https://github.com/JsLth)





## Your turn

- Let's check our the survey results...

1. Name?
2. Affiliation? Country?
3. What do you want to use Shiny for? (or research questions?)

<!--
Below take too long for short workshop
4. Previous experience with R?
5. Why are you interested in this
workshop?
-->




## Contact & Outline & Dates
- **Important**: 1st time we teach workshop/material
- **Course outline/content/dates:** (see toc on the left) 
  - Day 1: Introduction to Shiny
    - 13:00 - 14:00: Welcome and Introduction (1)
    - 14:00 - 15:00: Introduction (2)
    - 15:00 - 15:30: Coffee break
    - 15:30 - 16:30: User Interface (UI): Designing the front end (1)
    - 16:30 - 17:30: User Interface (UI): Designing the front end (2) [**Introduction tab**]
  - Day 2: Reactive programming
    - 13:00 - 14:00: Server: Reactive programming (1)
    - 14:00 - 15:00: Server: Reactive programming (2)  [**Tabulate data tab**]
    - 15:00 - 15:30: Coffee break
    - 15:30 - 16:30: Modelling and visualizing data (1)
    - 16:30 - 17:30: Modelling and visualizing data (2) [**Modelling data tab**]
  - Day 3: Develop your app
    - 13:00 - 14:00: Mapping data & advanced visualization (1)
    - 14:00 - 15:00: Mapping data & advanced visualization (2)  [**Mapping data tab**]
    - 15:00 - 15:30: Coffee break
    - 15:30 - 16:30: Theming & styling 
    - 16:30 - 17:30: Strengthen & Deploy






## Script & material {#material}
- **Literature**: See syllabus.
- **Website/script**: https://bookdown.org/paul/shiny_workshop/
	- Find it: Google "shiny paul jonas"
	- Document = slides  + script (Zoom in/out with `STRG + mousewheel`)
	- **Code**: can all be found in the script
	- **Data**: can usually be downloaded over links in the script. If not we'll share the files.
	- Full screen: F11
	- **Navigation**: TOCs on left and right
	- Search document (upper left)
	- Document generated with [quarto](https://quarto.org/)
- **Motivation**: Have a go-to script for participants (and ourselves!)
- **Content**: Mixture of theory, lab sessions, exercises and pure code examples for discussion




## Strategy & Goals
- **Strategy**: From the simple to the complex, slowly building up a complex Shiny app that includes various aspects (tabulate data, modelling and descriptive graphs, mapping)

- **Goals**: By the end of the course participants will:
	- know what the structure of a Shiny application looks like
	- understand the basics of reactive programming for interactive data analysis and visualization
	- be comfortable to use R Shiny to build their own interactive applications
	- have learned about different ways to launch their Shiny application






  






## Online vs. offline

* **Negative**
    + Screen fatigue
    + Can't run around to check your code
    + Less engaging, less social
    + Voice
    + Screen sharing &less screen space than classroom 
* **Positive**
    + We see the Shiny app how its mostly consumed
    ;-) (on a screen)
        + Remember: "How is your Shiny app consumed (smartphone)?"
    + Participation from everywhere
* **Rule(s)**: Please keep your camera online if possible!
  + Distracting animals/children/partners are a welcome distraction!
  + Yawning, leaving, looking bored etc. allowed!
  + Use a virtual background if you like!
  







  
    
## Recommended readings

- **Important**: Our workshop does not require any prior reading. 
- However, our schedule is primarily based on two textbooks which we generally recommend for further reading (see references on website):
	- @wickham2021mastering: Mastering Shiny: Build Interactive Apps, Reports, and Dashboards Powered by R. Accessible online at: https://mastering-shiny.org/. 
	- Fay, Colin, Rochette, Sébastien, Guyader, Vincent, and Girard, Cervan (2022): Engineering Production-Grade Shiny Apps. Accessible online at: https://engineering-shiny.org/.













## Software we will use
- Open-source software! (Q: Why?)
- R [@rcoreteam2023]^[Creators: [Core contributors](https://www.r-project.org/contributors.html)  and thousands of package authors.]
  - only viable competitor is Python
  - Install the necessary packages using the code below.

```{r workshop-1, echo=FALSE, comment=NA, cache=FALSE}
options(width = 300)
cat(packages)
```

- Shiny [@chang2020; @wickham2021mastering]
- Ggplot2^[Creators: https://github.com/tidyverse/ggplot2] [@wickham2016]
* Plotly^[Creators: https://github.com/plotly/plotly.js; https://github.com/ropensci/plotly] [@sievert2020]
* Note: Ideally cite the software you use in your research especially when it is open-source (e.g., run `citation("ggplot2")`)






## Helpful resources

1. [Shiny user showcase](https://shiny.posit.co/r/gallery/#user-showcase) 
2. [Shiny feature demos](https://shiny.posit.co/r/gallery/#feature-demos)
3. [Shiny widget gallery](https://shiny.posit.co/r/gallery/widgets/widget-gallery/)
4. [R Shiny for Enterprise Application Development: 6 Lessons for Building Shiny Apps](https://appsilon.com/why-you-should-use-r-shiny-for-enterprise-application-development/#anchor-7)
5. [Shiny apps by Dimiter Toshkov](http://www.dimiter.eu/Visualizations.html)
