---
title: "About this workshop"
---


```{r functions, echo=FALSE}
# Generate package chunk
   file_names <- stringr::str_subset(dir(), "qmd")
  packages_all <- NULL
  for(i in file_names){
  	#print(i)
  	
  lines_text <- readLines(i) 
  packages <- gsub("library\\(|\\)", "",  
                  unlist(stringr::str_extract_all(lines_text,  
                                         "library\\([a-zA-z0-9]*\\)|p_load\\([a-zA-z0-9]*\\)"))) 
  packages <- packages[packages!="pacman"] 
  packages <- packages[packages!="p_load("] 
  packages_all <- c(packages_all, packages)
  
  }
  packages <- paste("# install.packages('pacman')", "library(pacman)", "p_load('", paste(unique(packages_all), collapse="', '"), "')",sep="") 
  packages <- stringr::str_wrap(packages, width = 80) 
  packages <- gsub("install.packages\\('pacman'\\)", "install.packages\\('pacman'\\)\n", packages) 
  packages <- gsub("library\\(pacman\\)", "library\\(pacman\\)\n", packages)
```







## About us (Paul & Jonas)
- Paul
	- Currently [MZES](https://www.mzes.uni-mannheim.de/d7/en) Research fellow (University of Mannheim)
	- Previously... University of Bern (PhD, 2015), European University Institute (Postdoc, 15'-17')
	- My research interests: political sociology & methodology [see [Google Scholar](https://scholar.google.com/citations?user=hw-b-MkAAAAJ&hl=en&oi=ao))
	- Started R around 2009, first shiny apps ~2015 for teaching purposes
	- Contact: `mail@paulcbauer.de`; [www.paulcbauer.de](https://sites.google.com/paulcbauer.de/web/); [Twitter](https://twitter.com/p_c_bauer); [Github](https://github.com/paulcbauer)
- Jonas
	- Currently a research fellow at GESIS' survey data curation (SDC) department
	- Constantly learning about R and Shiny since 2020
	- Interested in social geography, spatial analysis and computational text analysis
	- Contact: `jonas.lieth@gesis.org`; GitHub: [JsLth](https://github.com/JsLth)





## Your turn

- Let's check our the survey results...

1. Name?
2. Affiliation? Country?
3. Research interests?
4. Previous experience with R?
5. Why are you interested in this
workshop?





## Workshop: Contact & Outline & Dates

- Framework
  + GESIS – Leibniz Institute for the Social Sciences
  + [Gesis Training Courses](https://training.gesis.org/?site=pOverview&cat=Workshop)
- **Dates**, **Contact** and **Outline:** See syllabus (received link via email).
  + **I hope you are fine with the schedule?**
    + Whenever you need a break let us know!




## Workshop: Script & material {#material}
* **Literature**: See syllabus.
* **Website/script**: https://bookdown.org/paul/shiny_workshop/
	- Find it: Google "shiny paul jonas"
	- Document = slides  + script (Zoom in/out with `STRG + mousewheel`)
	- **Code**: can all be found in the script
	- **Data**: can usually be downloaded over links in the script
	- Full screen: F11
	- TOCs on left and right
	- Search document (upper left)
	- Document generated with [quarto](https://quarto.org/)
* **Motivation**: Have a go-to script for participants (and ourselves!)
* Mixture of theory, lab sessions, exercises and pure code examples for discussion




## Workshop: Goals

- By the end of the course participants will:
	- understand the basics of reactive programming for interactive data analysis and visualization
	- know what the structure of a Shiny application looks like
	- be comfortable to use R Shiny to build their own interactive applications
	- have learned about different ways to launch their Shiny application


## Workshop: Structure & stategy
* See toc on the left
* Strategy: From the simple to the complex
	- Slowly building up a complex Shiny app that includes various aspects (data, descriptive statistics, modelling, map)
* Let's have a [look at the app]!




## Workshop: Online vs. offline

* **Negative**
    + Screen fatigue
    + Can't run around to check your code
    + Less engaging, less social
    + Voice
    + Screen sharing &less screen space than classroom 
* **Positive**
    + We see the graphs where they are mostly consumed
    ;-) (on a screen)
        + Remember: "How is your graph consumed (smartphone)?"
    + Participation from everywhere
* **Rule(s)**: Please keep your camera online!
  + Distracting animals/children/partners are a welcome distraction!
  + Yawning, leaving, looking bored etc. allowed!
  + Use a virtual background if you like!
  







  
    
## Workshop: Recommended readings

- Important: The course does not require any prior reading. 
- However, our schedule is primarily based on two textbooks which we generally recommend for further reading:
	- @wickham2021mastering: Mastering Shiny: Build Interactive Apps, Reports, and Dashboards Powered by R. Accessible online at: https://mastering-shiny.org/. 
	- Fay, Colin, Rochette, Sébastien, Guyader, Vincent, and Girard, Cervan (2022): Engineering Production-Grade Shiny Apps. Accessible online at: https://engineering-shiny.org/.













## Workshop: Software we will use

* What software will we use?
* Open-source software!
		+ Q: What speaks for open-source software?

* R (only viable competitor is Python)^[Creators: [Core contributors](https://www.r-project.org/contributors.html) and thousands of package authors.]
    + Install the necessary packages using the code below.

```{r workshop-1, echo=FALSE, comment=NA}
options(width = 300)
cat(packages)
```

* Shiny
	- ...
	
* Ggplot2^[Creators: https://github.com/tidyverse/ggplot2]
* Plotly^[Creators: https://github.com/plotly/plotly.js; https://github.com/ropensci/plotly]
* Note: Ideally cite the software you use in your research especially when it is open-source (see `citation("ggplot2")`)






## Helpful resources

1. Shiny user showcase: https://shiny.posit.co/r/gallery/#user-showcase
2. Shiny feature demos: https://shiny.posit.co/r/gallery/#feature-demos

