---
title: "Build your own app"
---

```{r, include=FALSE}
library(haven)
library(sf)
```

-   Learning outcomes:
    -   Become proficient in designing and developing a Shiny app from start to finish
    -   Develop and sketch out ideas about interactive visualization as Shiny app prototypes
    -   Acquire a holistic understanding of the Shiny development workflow

# Schedule for today

| Time          | Topic                                                |
|---------------|------------------------------------------------------|
| 10:00 - 11:00 | **Good practices when building your own app**        |
| 11:00 - 11:15 | Coffee break                                         |
| 11:15 - 12:00 | **Build your own app: Design and UI**                |
| 12:00 - 13:00 | Lunch break                                          |
| 13:00 - 14:15 | **Build your own app: Construct the reactive graph** |
| 14:15 - 14:30 | Coffee break                                         |
| 14:30 - 15:30 | **Build your own app: Strengthen & deploy**          |
| 15:30 - 16:00 | **Exhibit: Show off your new Shiny app**             |

This schedule is mostly a **suggestion**! Depending on the type of app you are building, some of these tasks require more or less time.

# Data

-   To build your Shiny app, we suggest one of the following datasets

## Bring your own dataset

-   If you already have a dataset that you want to build an app around, feel free to do this today!

## European social survey - wave 9

-   We used this dataset alongside the workshop
-   You can find
-   It contains data on political trust, political spectrum, demographics, internet use, happiness, and income
-   The dataset contains `haven` labels which help understand the scales of each variables
-   Base information:
    -   Region: Europe
    -   Time: 2018
    -   Observations: 49,519
    -   Variables: 14
-   How to read the data:

```{r, eval=FALSE}
library(haven) # the haven package has to be attached
ess <- readRDS("ess_trust.rds")
ess
```

```{r, echo=FALSE}
readRDS("data/ess_trust.rds")
```

-   Datamap:

| Column              | Description                                                                                                                                                                                     |
|-------------|----------------------------------------------------------|
| `idno`              | Respondent ID                                                                                                                                                                                   |
| `country`           | Country ID                                                                                                                                                                                      |
| `internet_use`      | People can use the internet on different devices such as computers, tablets and smartphones. How often do you use the internet on these or any other devices, whether for work or personal use? |
| `trust_parliament`  | How much do you personally trust the national parliament?                                                                                                                                       |
| `trust_legal`       | How much do you personally trust the legal system?                                                                                                                                              |
| `trust_police`      | How much do you personally trust the police?                                                                                                                                                    |
| `trust_politicians` | How much do you personally trust politicians?                                                                                                                                                   |
| `trust_parties`     | How much do you personal trust political parties?                                                                                                                                               |
| `trust_eu`          | How much do you personally trust the European parliament?                                                                                                                                       |
| `trust_un`          | How much do you personally trust the United Nations?                                                                                                                                            |
| `left_right`        | In politics people sometimes talk of 'left' and 'right'. Where would you place yourself on this scale, where 0 means the left and 10 means the right?                                           |
| `happiness`         | Taking all things together, how happy would you say you are?                                                                                                                                    |
| `age`               | Age of the respondent                                                                                                                                                                           |
| `income_feeling`    | Which of the descriptions on this card comes closest to how you feel about your household's income nowadays?                                                                                    |

## Guerry dataset

-   This dataset stems from the "Essay on the Moral Statistics of France" from 1830, written by André-Michel Guerry
-   It features a number of indicators of the so-called "moral statistics", including crime, literacy, wealth, donations, and more
-   The dataset is an `sf` dataset containing department polygons of 19th century France – it may thus be convenient for creating an app using `leaflet` maps
-   Base information:
    -   Region: France
    -   Time: 1815-1834
    -   Observations: 85
    -   Variables: 27
    -   Source: <https://sfdep.josiahparry.com/reference/guerry>
-   How to read the data:

```{r}
guerry <- readRDS(url("https://github.com/paulcbauer/shiny_workshop/raw/shiny_workshop_2024/data/guerry.rds"))
guerry
```

-   Datamap:

| Column            | Description                                                                                            |
|---------------|--------------------------------------------------------|
| `dept`            | Department ID                                                                                          |
| `region`          | Region of France (‘N’=’North’, ‘S’=’South’, ‘E’=’East’, ‘W’=’West’, ‘C’=’Central’)                     |
| `department`      | Department name                                                                                        |
| `crime_pers`      | Population per crime against persons                                                                   |
| `crime_prop`      | Population per crime against property                                                                  |
| `literacy`        | Percent of military conscripts who can read and write                                                  |
| `donations`       | Number of donations to the poor                                                                        |
| `infants`         | Population per illegitimate birth                                                                      |
| `suicides`        | Population per suicide                                                                                 |
| `main_city`       | Size of the principal city (‘1:Sm’, ‘2:Med’, ‘3:Lg’)                                                   |
| `wealth`          | Per capita tax on personal property                                                                    |
| `commerce`        | Rank of the number of patents per population                                                           |
| `clergy`          | Rank of the number of Catholic priests in active service population                                    |
| `crime_parents`   | Rank of the ratio of crimes against parents to all crimes                                              |
| `infanticide`     | Number of infanticides per population                                                                  |
| `donation_clergy` | Number of bequests and donations to the clergy per population                                          |
| `lottery`         | Per capita wager on Royal Lottery                                                                      |
| `desertion`       | Number of soldiers accused of desertion minus the deficit produced by insuffiency of available billets |
| `instruction`     | Inverse literacy                                                                                       |
| `prostitutes`     | Number of prostitutes registered in Paris                                                              |
| `distance`        | Distance to paris in km                                                                                |
| `area`            | Area in km²                                                                                            |
| `pop1831`         | Population in 1831, in thousands                                                                       |

# Exercises

::: callout-note
#### Exercise 1: 11:15 - 12:00

Taking one of the datasets introduced above – ESS, Guerry, or your own dataset, reflect on a possible Shiny app that you can build using the dataset. Create a sketch and design a prototype of the user interface based on this sketch.
:::

::: callout-note
#### Exercise 2: 13:00 - 14:15

Based on the prototype developed in exercise 1, sketch a reactive graph. Try to implement this reactive system in the server function of your app.
:::

::: callout-note
#### Exercise 3: 14:30 - 15:30

Deploy the app created in the last exercise using either <https://www.shinyapps.io/> or `shinylive`.
:::
